---
output: github_document
---

# Visual regression testing in the `tinytest` framework for `R`

<!-- badges: start -->
[![R-CMD-check](https://github.com/vincentarelbundock/tinyviztest/workflows/R-CMD-check/badge.svg)](https://github.com/vincentarelbundock/tinyviztest/actions)
<!-- badges: end -->

[`tinytest` is a wonderful package for `R`](https://cran.r-project.org/package=tinytest) created by Mark van der Loo. It is a "lightweight, no-dependency, but full-featured package for unit testing in `R`."

`tinyviztest` extends `tinytest` with an expectation function called `expect_vdiff()`. This function can be used to:

1. Take snapshots of known "target" plots.
2. Test if the "current" plot matches the target.

`tinyviztest` supports both `ggplot2` and base `R` plots.

Under the hood, `tinyviztest` uses [the `gdiff` package by Paul Murrell](https://cran.r-project.org/package=gdiff) to compare plots.

## Installation

`tinyviztest` is not yet available on CRAN. It can be installed from Github:

```{r, eval = FALSE}
library(remotes)
install_github("vincentarelbundock/tinyviztest")
```

## How to use `tinyviztest` in an `R` package

The `tinyviztest` repository itself acts as an example for how to setup up visual tests:

* https://github.com/vincentarelbundock/tinyviztest/blob/main/inst/tinytest/test-basic.R
* https://github.com/vincentarelbundock/tinyviztest/blob/main/tests/tinytest.R

## `ggplot2`

The first time we run `expect_vdiff()`, the test fails, and a target plot is saved in a folder called `_tinyviztest`

```{r}
library(ggplot2)
library(tinytest)
using(tinyviztest)

p1 <- ggplot(mtcars, aes(mpg, wt)) + geom_point()
p2 <- ggplot(mtcars, aes(mpg, hp)) + geom_point()

expect_vdiff(p1, name = "ggplot2 example")
```

The second time we run `expect_vdiff()` the test passes:

```{r}
expect_vdiff(p1, name = "ggplot2 example")
```

If we use a different plot, the test fails:

```{r}
expect_vdiff(p2, name = "ggplot2 example")
```

We can overwrite the target plot with an argument:

```{r}
expect_vdiff(p2, name = "ggplot2 example", overwrite = TRUE)

expect_vdiff(p2, name = "ggplot2 example")
```

## Base `R` graphics

To test a Base `R` plot, we supply a function which prints the plot:

```{r}
p1 <- function() plot(mtcars$hp, mtcars$wt)
p2 <- function() plot(mtcars$hp, mtcars$mpg)

expect_vdiff(p1, name = "base")

expect_vdiff(p1, name = "base")

expect_vdiff(p2, name = "base")
```

